### 2021-06-11 draft_glossa-psych/glossa-psych_AM-ID_manuscript.tex
final manuscript

% !TeX document-id = {6d544fe2-4255-4108-8022-d809ff1d2c53}
% TeX program = xelatexmk %doesn't work, no xelatexmk option under 'build'. Changed default engine to xelatex (options-->configure-->build-->default build-->xelatex)
% see http://info.semprag.org/basics for a full description of this template
\documentclass[charis,linguex]{glossa}
% possible options: 
% [times] for Times font (default if no option is chosen)
% [cm] for Computer Modern font 
% [lucida] for Lucida font (not freely available)
% [brill] open type font, freely downloadable for non-commercial use from http://www.brill.com/about/brill-fonts; requires xetex
% [charis] for CharisSIL font, freely downloadable from http://software.sil.org/charis/
% for the Brill an CharisSIL fonts, you have to use the XeLatex typesetting engine (not pdfLatex)
% for headings, tables, captions, etc., Fira Sans is used: https://www.fontsquirrel.com/fonts/fira-sans
% [biblatex] for using biblatex (the default is natbib, do not load the natbib package in this file, it is loaded automatically via the document class glossa.cls)
% [linguex] loads the linguex example package
% !! a note on the use of linguex: in glossed examples, the third line of the example (the translation) needs to be prefixed with \glt. This is to allow a first line with the name of the language and the source of the example. See example (2) in the text for an illustration.
% !! a note on the use of bibtex: for PhD dissertations to typeset correctly in the references list, the Address field needs to contain the city (for US cities in the format "Santa Cruz, CA")

%\addbibresource{sample.bib} 
% the above line is for use with biblatex
% replace this by the name of your bib-file (extension .bib is required)
% comment out if you use natbib/bibtex

%for natbib: replace \cite with \citep (parenthetical); \citea with \cite (in-text)
\usepackage{hyperref}
\let\B\relax %to resolve a conflict in the definition of these commands between xyling and xunicode (the latter called by fontspec, called by charis)
\let\T\relax
\usepackage{xyling} %for trees; the use of xyling with the CharisSIL font produces poor results in the branches. This problem does not arise with the packages qtree or forest.
\usepackage[linguistics]{forest} %for nice trees!

\usepackage{tipa}
\newcommand{\nt}[1]{\textipa{[#1]}} % narrow transcription
\newcommand{\wt}[1]{\textipa{/#1/}} % wide transcription

%\graphicspath{{images/}}

% \pdf* commands provide metadata for the PDF output. ASCII characters only!
%\pdfauthor{Auromita Mitra, Indranil Dutta}
%\pdftitle{Mixed language processing increases cross-language phonetic transfer in Bengali-English Bilinguals}
\pdfkeywords{phonetic transfer, vowel formants, Bengali, English}

\title{Mixed language processing increases cross-language phonetic transfer in Bengali-English bilinguals}
% Optional short title inside square brackets, for the running headers.

% \author[Auromita \& Indranil]% short form of the author names for the running header. If no short author is given, no authors print in the headers.
% {%as many authors as you like, each separated by \AND.
%   \spauthor{Auromita Mitra\\ 
%   \institute{The EFL University}\\
%   \small{%105, Bd. Raspail, 75005 Paris\\
%   auromita.mitra@gmail.com}
%   }
%   \AND
%   \spauthor{Indranil Dutta \\
%   \institute{Jadavpur University}\\
%   \small{%Warmoesberg 26, 1000 Brussel\\
%   indranildutta.lnl@jadavpuruniversity.in}
%   }%
% }
%\usepackage{gb4e}

\begin{document}

\sffamily
\maketitle

Word count: 9263

\begin{abstract}
This study investigates the phonetic outcome of mixed-language production in a group of highly proficient bilingual speakers of Bengali and English, to better understand the nature of cross-language phonetic interaction. Existing research has largely focused on a small set of phonologically related languages and used voice onset time (VOT) to measure phonetic transfer. Results indicate that such interaction is mediated by both language-specific features and linguistic experience, highlighting the need for diversification in the range of languages and populations studied. We compare productions from a yet-unstudied population in two conditions-- unilingual L2 (English) and mixed (Bengali and English). The spectral properties of two L2 vowel categories— \nt{\ae} (common to L1 and L2) and \nt{2} (absent in L1) are examined as sites for transfer. To understand how this unfolds dynamically, we measure F1 and F2 at five points over the course of articulating the vowel. We find dynamic and asymmetrical changes in both vowels during mixed production, with \nt{2} shifting towards a related L1 category \nt{a:}. We interpret this as a temporary increase in cross-language phonetic influence during mixed-language use, leading to a shift towards L1 norms. Mixed language is elicited in two switching paradigms: cued picture-naming and code-switching, to assess if differences between the paradigms independently influence the outcome of phonetic interaction. Results reveal no difference in transfer patterns across the two paradigms. We discuss these findings in light of recent proposals about asymmetries in short-term phonetic interaction and the postulated role of connected speech in introducing discursive factors to transfer studies. 
\end{abstract}

\begin{keywords}
 phonetic transfer; vowel quality; bilingual speech production; L1-L2 interaction; Indian English; Bengali 
\end{keywords}

\rmfamily

\section{Phonetic transfer}

Bi/multilingual speakers can distinguish between the phonetic norms of their languages and maintain separate sound categories for each language \citep{caramazza1973acquisition,macleod2010impact,bosch2003simultaneous}. However, these categories are not autonomous-- they influence each other across languages in both perception and production \citep[e.g.][]{flege1995second,fowler2008cross,flege2002assessing}, and the nature of such interaction provides crucial insights into how language ``systems" are cognitively represented and processed. Based on the productions of proficient bilingual speakers of Bengali and English, this study reports that cross-language phonetic influence temporarily increases during mixed-language use.\\%Given that a majority of the world's population is multilingual, this is a central concern for any realistic theory of language.
%show cross-language-- rephrase

%keep just the types here, and put the next 2 lines in the causes&effects section?
Cross-language influence at the level of sounds can be studied in broadly two kinds of conditions:
(i) While a bilingual speaker is operating in any one of their languages, by comparing bilingual speech to monolingual norms \citep[e.g.][]{guion2003vowel,caramazza1973acquisition,flege1987production}; %These studies often compare the speech of bilinguals to monolingual norms, and view the influence as changes to long-term memory representations as a result of acquiring an L2 \citep{guion2003vowel,caramazza1973acquisition,flege1987production}; 
(ii) When both languages of a bilingual speaker are co-activated, by comparing speakers' mixed-language speech to their own norms while using a single language \citep[e.g.][]{grosjean1994going, bullock2009trying,elias2017effects, simonet2014phonetic}. %These compare productions during mixed-language use to participants' own productions while using a single language. The resulting influence is variously thought to involve short-term memory, online processing costs, language mode, and context-awareness. 
The present study is concerned with the latter. In production, this kind of influence has been variously termed \textit{transfer, drift, accommodation,} and \textit{interference}-- we use these interchangeably here to indicate any interaction between two sets of phonetic norms.\\

Existing research on the phonetic effects of mixed-language production has largely focused on a limited set of phonologically related languages. A majority of these studies use temporal properties of consonants (in particular, voice onset time or VOT) to measure transfer. However, the reported results vary greatly across studies, and appear to be contingent upon both language-specific features and language experience of the participants. This emphasizes the importance of considering data from a wider variety of
%elaborate on 'great variation'-- differences between language pairs, plus sociolinguistic situation (effects of language experience etc)-- clearer. 'Types of features'--separate line?- Brief mention - discursive factors in VOT. 
populations, language pairs,  and phonetic features in order to make meaningful generalizations. Widespread multilingualism in the Indian subcontinent suggests that phonetic
%line--rephrase
behavior in these populations can be particularly valuable towards understanding the nature of such cross-language interactions, as they are likely to reflect real-world experience with mixed-language processing. However, there is no work yet on short-term phonetic transfer in the Indian subcontinent, or in any Indo-Aryan language. \\

The present study examines phonetic transfer between Bengali and English in a group of highly proficient bilingual speakers in India. We measure spectral properties: F1 and F2, of two English vowels to ascertain if L1 influence on L2 increases during mixed-language use, relative to a participant's unilingual baseline production of L2. Mixed-language data is elicited in two switching paradigms: cued picture-naming and code-switching. We compare these results to assess if differences between the paradigms independently influence the outcome of phonetic interaction. The results demonstrate a dynamic effect of transfer on L2 vowels during mixed-language production, but no difference in the extent of transfer between the two paradigms. We discuss these findings in light of recent proposals about asymmetries in short-term phonetic interaction and the role of connected speech in introducing discursive factors to transfer studies.\\

The rest of this section is structured as follows: \ref{bengali_english_in_india} provides background information on Bengali and Indian English focusing on vowel systems. \ref{causes} discusses two potential sources of transfer during language processing-- global co-activation of multiple languages vs the local act of switching between languages. Given the highly multilingual setting that characterizes this population, we note that the former is expected to be a constant feature and therefore any observed phonetic interaction is better understood as resulting from the latter. The design of the present study is motivated on the basis of this discussion. A logical consequence of this model is that such phonetic interaction must be highly localized. \ref{duration} elaborates on this by reviewing literature on the duration and ``reversiblity" of transfer effects. A recurring pattern that emerges from existing research on short-term phonetic transfer is that both languages of the bilingual speaker are not equally affected, although the nature of these reported differences varies greatly across studies. In \ref{asymmetries}, we review some of these findings, focusing on proposed sources of asymmetries, in order to highlight the factors that mediate such phonetic interaction (or fail to do so). We argue that these results point to an urgent need for expanding the range of language pairs and sound categories examined in order to make meaningful generalizations about the nature of short-term phonetic transfer. Following from this, \ref{asymmetry between sounds} discusses reported asymmetries between different sound categories of a single language, and develops the hypotheses of the present study on the basis of these. In \ref{paradigms}, we motivate a secondary aim of this study: to verify the postulated differences between two paradigms for eliciting short-term phonetic interaction in an experimental setup, namely cued picture-naming and code-switching. \ref{questions_and_hypotheses} summarizes the research questions and hypotheses.

\subsection{Bengali and English in India} \label{bengali_english_in_india}

\subsubsection{Demography}

Bengali (also, Bangla) is an Indo-Aryan language primarily spoken in India and Bangladesh. In India, more than 97 million people speak Bengali as a first language (Census of India, 2011), mostly in the state of West Bengal. A majority of this population also speaks other additional languages.

Indian English (IE) refers to the variety of English that has developed in the Indian subcontinent. In India, it is spoken as an L2 by 129 million people (Census of India, 2011).  English is one of the two official languages, used in education, law, media, as a \emph{lingua franca} mainly for an educated elite in most metropolitan regions, and carries a high prestige value \citep{pandey201517, kachru1981english, tollefson2014language, kachru1983indianization}.
%citations 
Recent research comparing the segmental and suprasegmental properties of IE spoken in different parts of the country report many commonalities (see \cite{sirsa2013effects} for a review). This suggests that in spite of the varied L1s of its speakers, IE has a target phonology that is distinct from any of these, as well as from other native varieties of English. Thus, it is fruitful to think of regional variations as resulting from L1-influence on a common underlying target. Excepting a small minority of specific communities \citep{pandey201517, wells1982accents, coelho1997anglo}, there are very few L1 speakers of IE.\\

%%%%%%%%%%%%%%%%%%%%%%%do something%%%%%%%%%%%%%%%
Considering the language usage patterns suggested by this demography, the present study focuses specifically on short-term phonetic interaction during mixed-language use, because:

\begin{enumerate}[label=(\roman*)]
	\item In multilingual populations, long-term representations of sound categories are expected to be affected by multiple languages. 
	\item Given that examples of an IE phonology without L1 ``influence" are very rare, it is more meaningful to think of cross-language transfer in L2 as relative to a speaker's own production in a given baseline condition.
\end{enumerate}

The socio-linguistic facts about English in India also suggest that for the present population, mixed-language processing of English and an L1 is an ecologically valid paradigm that is likely to be a part of everyday language experience. Thus, data from this population is valuable to the understanding of short-term phonetic interaction.

\subsubsection{Vowel systems} \label{vowel systems}
The vowel inventory of Western Bangla (the variety spoken by the participants in this study) consists of \nt{i, e, \ae, a, O, o, u}, and their nasalized counterparts \citep{garry2001facts}. Note that there is no mid-central vowel category.

The vowel system of IE contains the monopthongs \nt{I, i, E, e, \ae, @/2, a:, O, o, U, u}, represented by the lexical set KIT, FLEECE, DRESS, FACE, and TRAP, STRUT, PALM, LOT/CLOTH, GOAT, FOOT, GOOSE \citep{wells1982accents, masica1972sound}. %Their distribution in the F1XF2 space is shown in (insert table 1.7, Pandey). 
A single mid-central vowel corresponds to the categories \nt{2,@,3:}, which are treated as distinct in many native varieties of English \citep{nihalani1979indian,wells1982accents,hickey2005legacies,bansal1969intelligibility}. Since the English items used in this study are traditionally transcribed with \nt{2}, we use this symbol to indicate the mid-central vowel throughout.

\subsection{What causes transfer and what does it affect?}\label{causes}

Existing research has distinguished changes in category \textit{representations} due to the acquisition of multiple sound systems (cf. Speech Learning Model \citep{flege1995second,flege2007language}; Perceptual Assimilation Model-L2 \citep{best2007nonnative}), from interaction during accessing, processing, or articulation of these categories (cf. transfer vs interference \citep{grosjean2012attempt}; competence vs performance interference \citep{paradis1993linguistic}). 
Given their transient nature, dynamic changes in production during mixed-language use are generally attributed to the latter, e.g. online processing costs \citep{olson2013bilingual,tsui2019impact,vsimavckova2015immediate}, language mode \citep{simonet2014phonetic}, context-awareness \citep{khattab2013phonetic}. What triggers this interaction? \cite{olson2016role} argues that while cross-language phonetic effects are largely measured at the point of language switch, it could have two potential sources:
\begin{enumerate}
	\item The local point of switch itself
	\item Global co-activation of two languages -- bilingual language mode \citep{grosjean1998studying} 
\end{enumerate}

A number of studies have specifically manipulated language mode, both in the presence and absence of switching. Overall, results suggest that: (i) In the absence of other manipulations, productions in a bilingual language mode show increased cross-language influence compared to a monolingual mode \citep{simonet2020increased,simonet2014phonetic}; (ii) However, language mode is not the sole source of influence during mixed language use -- studies comparing switched and non-switched tokens produced in the same test block (identical language mode) \citep{olson2016role,tsui2019impact} or spontaneous conversation \citep{piccinini2015voice}, have still reported a difference, suggesting that independently of mode, switching between languages triggers a local increase in cross-language transfer. (iii) How the two sources interact to influence the final outcome of transfer is not fully understood-- \cite{olson2016role} found no additive effects, \cite{olson2013bilingual} found the balanced language context to inhibit transfer compared to unbalanced contexts.  Other studies have not analyzed the two separately, eliciting switched tokens in a bilingual test block and non-switched tokens in a separate monolingual test block, separated by a few hours to days \citep{schwartz2015language, bullock2009trying,antoniou2011inter, elias2017effects,vsimavckova2015immediate,vsimavckova2018patterns}.\\

\cite{grosjean1998studying} suggests that various aspects of the communicative setting, including exposure to (spoken or written) stimuli in multiple languages and awareness of the interlocutor's being bilingual, could trigger a bilingual mode. The participants in the present study are in an environment which largely contains mixed-language input, multilingual interlocutors, and no stable ambient language. Therefore we expect that phonetic transfer during everyday language use, if any, takes place in a bilingual mode.  To preserve ecological validity, we elicited both unilingual and switched utterances in a bilingual language mode. Any observed differences in this paradigm would arguably result from interaction during online processing. \\

In a bilingual mode, both language systems are expected to be nearly equally accessible throughout the test block. Thus, a consistent difference between switched and non-switched tokens in such a paradigm would be possible only if the effects were highly localized-- if not, we should expect a gradual convergence over the course of the experiment. This is discussed in the next section.

\subsection{Duration of transfer effects} \label{duration}

Research comparing bilinguals speakers' phonological systems with monolinguals treats them as relatively stable over time. The difference from monolingual norms is interpreted as the cumulative result of cross-language influence over long periods \citep[e.g.][]{guion2003vowel, caramazza1973acquisition}. However, longitudinal studies %(citations) 
and between-subject comparisons of bilingual speakers with different durations of L2 exposure suggest that interaction between the sound systems is dynamic: over time, increasing exposure to an L2 can reduce effects of L1 influence, leading to a more fine-grained separation and native-like production of foreign contrasts. For example, \cite{bohn1992production} 
observed that approximately 7 years' difference in L2 exposure between groups having the same age of acquisition led to a significant difference in the amount of cross-language transfer. These have been interpreted as changes to long-term phonological representations.

Moreover, changes due to transfer are not unidirectional or irreversible. \cite{sancier1997gestural} first demonstrated that spending 2-5 months in an L1 or L2 environment causes productions in both languages to ``drift" towards the ambient language. This not only evidenced that cross-language interaction can be triggered in an order of months, but also that the effects can be reversed within a similar time range. A more recent study by \cite{tobin2017phonetic} reported comparable effects in an even shorter duration (2-4 weeks). Based on a short-term longitudinal study, \cite{chang2012rapid} demonstrates that over the course of the first five weeks of learning an L2, there is a gradual convergence of L1 towards L2. For some sounds and features (though not others, cf. sec.\ref{asymmetry between sounds}), transfer was additive over time. How long these effects last in the absence of regular L2 input was not tested. \\
% if effects are cumulative, then the difference between switched and non-switched tokes should be lesser in later test blocks (previous two paragraphs needed?)

The majority of studies on the phonetics of mixed-language use only analyze the switched (target) token. Thus, there are few direct measurements of the duration of short-term transfer effects. One study which measured this in a code-switching paradigm \citep{bullock2009trying} did not find any residual effects on the matrix language following a switch, suggesting that transfer during code-switching is localized. Indirect evidence in support of this comes from experiments which have elicited switched and non-switched tokens in the same test block and still reported differences between the two \citep[e.g.][]{tsui2019impact,olson2013bilingual}, suggesting that changes due to transfer are quickly ``reset"-- in an order of seconds.\\
Note that all the studies discussed above measure VOT, which is a temporal feature. We have no apriori reason to assume that these durations generalize to vowel quality. However, findings from sub-categorical phonetic shifts triggered by other factors (such as convergence towards an interlocutor) do evidence rapid shifts in vowel quality within comparable time-frames \citep[e.g.][]{pardo2010expressing,babel2010dialect,babel2012evidence}. Thus in the current study, we present switched and non-switched tokens randomly within the same test block to induce a bilingual mode. We expect the intervening words between two subsequent targets to undo any residual effects of transfer.

\subsection{Asymmetries between languages in extent and direction of transfer} \label{asymmetries}
A recurring pattern that emerges from existing research on short-term phonetic transfer is that the extent and patterns of phonetic shift are not equivalent across the two languages of the bilingual speaker. Studies vary greatly in the nature of differences that they report. The potential causes for such asymmetry are of interest because they point towards the factors that mediate cross-language phonetic interaction. In this section, we discuss two proposed sources of such asymmetry that are relevant to the present study:

\subsubsection{L1 vs L2 status} 
Flege's Speech Learning Model (SLM) \citeyearpar{flege1995second,flege2007language} posits that the sound categories of a bilingual speaker exist in a common phonological space, and therefore in principle, both L1 and L2 categories can influence one another. Thus, the patterns of cross-language transfer depend on how L2 phonemes are mapped in relation to the existing L1 categories. In non-switched production, this is evidenced through a pervasive L1 influence on non-native contrasts that are perceptually linked to an existing L1 contrast (cf. \textit{equivalence classification}-- \cite{flege1984limits,flege1987production}), and the observation that both L1 and L2 sound systems of bilinguals differ from those of corresponding monolingual speakers \citep[e.g.][]{guion2003vowel}. \\

In mixed-language production, the role of language status is less clear-- studies have reported unidirectional influence of L1 on L2 \citep{balukas2015spanish,antoniou2011inter,vsimavckova2015immediate,goldrick2014language}, L2 on L1 \citep{tsui2019impact,elias2017effects, olson2013bilingual}, bidirectional convergence \citep{bullock2009trying, olson2016role}, divergence \citep{bullock2009trying,vsimavckova2018patterns}, and no influence \citep{muldner2019phonetics,schwartz2015language}. Since most existing studies measure shifts in VOT, the observed asymmetries between languages have been variously explained either in terms of their L1 vs L2 status, or language-specific differences between long- and short-lag VOT languages. %(cf sec. \ref{sound systems}).

\cite{olson2013bilingual} first reported a unidirectional VOT shift of L1 towards L2 in two different groups -- native speakers of Spanish (short-lag) and English (long-lag), matched for proficiency and age of L2 acquisition. This established that beyond language-specific differences, the L1 vs L2 status of the language does mediate transfer. They interpret this asymmetry in terms of the Inhibitory Control Model (ICM) \citep{green1998mental}: to select a phonetic realization from one language, the other must be inhibited. Since L1 is the ``stronger" language, the amount of inhibition required on L1 while using an L2 is greater than the inhibition required on an L2 while using the L1. This greater initial inhibition on L1 means that switching into L1 incurs a greater ``switch cost" (more cross-language influence) than switching into L2. Lower switch cost results in a lack of visible cross-language transfer effects on L2. \cite{tsui2019impact} report comparable results, but only in participants who were not equally dominant in both languages. Balanced bilinguals did not demonstrate any transfer effects in VOT. They propose that this is because balanced bilinguals have better inhibitory control (low switch cost in both languages) due to greater experience with language switching.\\

Since differences between long-lag and short-lag languages independently affect the phonetic realization of VOT, focusing on other phonetic features as sites for transfer can avoid this conflation, and clarify the precise effect of language status. However, given the implications of the ICM, it is necessary to first establish that these L2 categories can indeed be affected by dynamic interference, particularly in proficient bilinguals. The few existing studies on vowel quality \citep{simonet2014phonetic,muldner2019phonetics,elias2017effects} or phonological processes \citep{simonet2020increased,schwartz2015language} have examined transfer effects on L1 categories. As far as we are aware, there is no work yet on vowels in L2. Thus, this study builds on the existing work by examining whether the L2 vowel quality of proficient bilinguals can be affected by mixed-language processing.

\subsubsection{Differences between sound systems of the languages} \label{sound systems} In addition to the cognitive factors discussed above, many researchers have attributed the observed asymmetries to language-internal factors. Specifically, a pattern that emerges from the body of work on VOT is that a shift in long-lag VOT towards the short-lag norms of the other language is much more consistent and systematic than the converse \citep{tobin2017phonetic, olson2016role,bullock2009trying,antoniou2011inter,chang2012rapid}. In general, short-lag VOTs seem to resist accommodative shift. \cite{bullock2009trying} suggest that because long-lag languages offer a greater range of acceptable VOTs, there is more ``room" for movement. In contrast, shift in short-lag languages could risk the loss of a phonological contrast (VOT being the primary cue for voicing contrast in these languages).  This is a language-specific phonological constraint on transfer.
In a similar vein, \cite{antoniou2011inter} observe that while the VOT of English stops shifted towards Greek (a short-lag language), the reverse was seen only in a specific subset of sounds-- word-medial non-nasalized stops. They attribute this to the low frequency of these sounds in Greek, making them less ``stable", and thus more amenable to cross-language influence. This suggests that the distribution of sounds in a language could affect the kind of transfer patterns observed.\\

These results highlight that beyond interaction during the cognitive processing of sounds, phonetic transfer is ultimately a linguistic phenomenon, and thus subject to language-specific phonological constraints. Therefore, to make meaningful generalizations, it is imperative to consider data from a variety of language pairs. Existing research on phonetic transfer largely revolves around a limited set of phonologically related languages. Thus, the present study extends the scope of this research to a new pair of languages--- Indian English and Bengali.

Bengali has a four-way laryngeal contrast. Thus, VOT does not encode a binary voiced-voiceless distinction, and moreover is not the only relevant cue for the corresponding phonological categories \citep{dmitrieva2020acoustic}. This means that any effect of L1 Bengali on L2 English VOT is likely to be mediated by additional phonological factors. Thus, avoiding any potential confounds was another reason we chose to focus on vowel quality as a target for transfer.

\subsection{Asymmetries between sounds}\label{asymmetry between sounds}

Asymmetries in the extent and patterns of transfer have not only been observed between languages, but also between different sounds/features of a language, in both long-term and transient interactions. Studies which have examined multiple sound categories have found that interactions between individual sounds pairs do not necessarily reflect the overall pattern of global (system-wide) shift \citep[e.g.][]{chang2012rapid,elias2017effects}, suggesting that \textit{extent of transfer} cannot treated as an atomic measure. In light of the discussion about VOT in section \ref{sound systems}, this is not surprising --- if a general linguistic principle of ``room for movement" constrains transfer, then we should expect it to apply to individual sounds too. Once again, this emphasizes the importance of examining a wider range of sound contrasts.\\

In the present study, we focus on two vowel categories in Indian English-- the mid-central vowel \nt{2} and the low front vowel \nt{\ae}. There are two plausible sources of asymmetry between these:
\begin{enumerate}
	\item Position in the IE vowel space: compared to \nt{\ae}, \nt{2} exists in a part of the vowel space that has a lower vowel density. This affords a greater latitude for movement without risking the loss of a contrast, particularly in the vowel height (F1) dimension. Thus, considering purely phonological constraints on IE, we expect that a greater degree of shift is possible in \nt{2} compared to \nt{\ae}. Any movement in \nt{\ae} is expected to be primarily in the backness (F2) dimension.
	\item Target of transfer: findings from existing research on cross-language phonetic interaction suggest that changes in production during mixed-language use are not random, but rather targeted with respect to categories in the other language. Thus, another source of asymmetry is the fact that the category \nt{2} is absent in Bengali, whereas \nt{\ae} is a common category across both languages (cf. sec.\ref{vowel systems}).
\end{enumerate}\\

%table

Flege's Speech Learning Model (SLM) \citeyearpar{flege1995second,flege2007language} posits that sound categories which are common across languages influence each other because they share a common acoustic-phonetic space. Thus, if Bengali and English differ in their canonical realizations of \nt{\ae}, then we should expect the English \nt{\ae} to shift towards the corresponding Bengali category in the mixed condition.

%Other categories can in principle be produced with native-like accuracy, showing little effect of transfer. Given that our participants are highly proficient bilinguals, and assuming that cross-language correspondences remain the same in representation and processing, this would lead us to expect more transfer in the common category, \nt{\ae}.

There is no obvious competing L1 category during the production of \nt{2}. However, it is unlikely that this should altogether preclude a shift in \nt{2}, since at least one existing study has reported transfer effects on a non-common vowel category in a comparable paradigm \citep{simonet2014phonetic}. Here, participants' Catalan \nt{O} shifted towards an acoustically close category \nt{o}, which is common to Catalan and Spanish, suggesting that movement might be towards a set of phonetic norms rather than a specific corresponding category.

We expect the vowel \nt{2} to shift towards an acoustically/perceptually close category in IE -- a probable candidate being the low vowel \nt{a:}, as anecdotal observations of Bengali-accented English suggest that the sounds are perceived as close by Bengali speakers. \\

In the present study, we do not elicit unilingual Bengali productions. To estimate average baseline differences in vowel quality between the relevant English and Bengali categories, we used Bengali production data from five female speakers from \cite{dutta2021}. Formant values were extracted from syllable-initial productions of \nt{\ae} and \nt{a:}, at 55\% into the vowel. Next, we compared these to our participants' productions in the unilingual English condition. Comparisons of Lobanov-normalized F1 and F2 suggest that the Bengali category \nt{\ae} is lower, and \nt{a:} lower and more fronted, than the English categories \nt{\ae} and \nt{2} respectively.

\subsection{A note on paradigms: connected speech or not?} \label{paradigms}

This section motivates a secondary aim of this study: to verify if the extent and patterns of phonetic interaction in bilingual participants are independently mediated by the choice of experimental paradigm. Studies of phonetic transfer during mixed-language use have largely employed spontaneous or scripted code-switching (CS) to elicit mixed productions. CS is defined as the use of multiple languages within a single utterance, i.e. in connected speech \citep{myers1993dueling}. Other paradigms which have been used include cued picture-naming, delayed repetition, interpreting across languages, reading word-lists etc. \cite{olson2013bilingual} argues that code-switching introduces discursive factors such as discourse planning and pragmatic considerations, which ultimately reflect patterns of language practice, rather than cognitive behavior. This finds support from studies which report transfer effects \textit{before} the switch point \citep{bullock2009trying}, or other evidence for planning such as hyperarticulation \citep{muldner2019phonetics} and interlocutor-awareness in spontaneous CS \citep{khattab2013phonetic}. To avoid such confounds, \cite{olson2013bilingual} endorses cued picture-naming (where pictures are named in rapid succession in either language, based on the given language cue) as an alternative paradigm to isolate purely phonetic effects in controlled setups. This observation has important implications for how the existing literature on short-term phonetic interaction is interpreted: if the discursive properties of a CS paradigm independently affect the phonetic outcome of mixed-language use, this calls to question the comparability of results across studies using these different paradigms. Identifying paradigm-specific differences is thus particularly relevant for this line of research, as generalizations often rely crucially on the assumption of such comparability. \\

Note that while there is evidence for planning in CS, precisely how this affects phonetic interaction cannot be predicted from the existing literature--- while \cite{bullock2009trying} found tokens immediately preceding the switch point to be phonetically distinct from the rest, one participant group converged towards the switch language (explained in terms of proficiency) while the other showed divergence (explained in terms of extra-linguistic factors). The participants in Muldner et al.'s \citeyearpar{muldner2019phonetics} study hyperarticulated switched tokens (indicating some degree of planning), and did not evidence a significant shift during mixed-language use. Moreover, recall that the findings of \cite{olson2016role} suggest that there is a limit on the extent of transfer-- multiple factors do not necessarily result in additive effects. Thus, the presence of additional factors in CS does not necessitate that this should alter the extent of observable transfer compared to any other paradigm. It is possible that the outcomes in the studies discussed above are better attributed to some other aspect of the experimental setup, rather than to the act of code-switching in particular.\\

The precise contribution of various (linguistic and extra-linguistic) factors and their interactions to the phonetics mixed-language use merits detailed investigation. In the present study, our focus is limited to verifying whether, in an experimental setting, the postulated discursive elements of a CS paradigm alter the extent of observable transfer compared to a paradigm that arguably lacks these--- cued picture-naming. We are less concerned with the exact nature of differences than with their comparability. While researchers have successfully employed each of these paradigms to induce phonetic interaction, comparison across studies is difficult since they also differ in other factors. We provide a starting point for such comparison by eliciting productions of the same tokens from the same set of participants, in both paradigms. A difference in results will suggest that the planning involved in connected speech affects transfer above and beyond the phonetic outcome of switching itself, which has implications that future work should consider.

\subsection{Research questions and hypotheses}\label{questions_and_hypotheses}
\begin{itemize}
	\item \textbf{L2 vowels}: Are L2 vowel categories of proficient bilingual speakers affected by dynamic phonetic transfer during mixed-language production, as evidenced by differences in the spectral properties of L2 vowels in switched and non-switched tokens?
	\item \textbf{Asymmetry}: If so, we expect an asymmetry between the vowels in the extent and direction of shift. Based on comparisons with Bengali vowels from \cite{dutta2021} we expect \nt{2} to be lowered and fronted in the mixed condition. \nt{\ae} is expected to show lowering, but to a lesser extent than \nt{2}, as it is in a part of the IE vowel space that has greater vowel density.
	\item \textbf{Paradigm}: Is there an observable difference between the extent of transfer in a cued picture-naming vs a code-switching paradigm? 
\end{itemize}

\section{Methodology} \label{methodology}
The experiment had the following factorial design: 10 participants* 20 target words* 2 conditions* 2 tasks* 4 iterations = 3200 target tokens.

\subsection{Participants}

10 Bengali-English bilingual speakers (male=5, age range 19 to 28), completed the tasks. All the participants speak other languages in addition to Bengali and English. At the time of recording, they were living in or around the EFL University campus in Hyderabad, which has a linguistically diverse population. Speakers were selected on the basis of responses to a Language Background Questionnaire to ensure comparable LSRW skills in L1 and L2 across participants, and were compensated for their time.

\subsection{Stimuli}
Twenty monosyllabic English words containing the vowels \nt{2} or \nt{ae} were selected as target words. To minimize coarticulatory effects on the vowel, the pre-vocalic consonants were all voiced plosives (\nt{b} or \nt{d}). 10 monosyllabic English words that do not contain the target vowel served as filler items. We avoided English words that are lexicalized loanwords in Bengali. %The list of words can be found in sec.\ref{appendix}.

\subsection{Procedure}
Participants were recorded individually in a sound attenuated room. %, using a ... microphone type/description?
Productions were recorded in Praat \citep{boersma2016praat} at a 44.1K sampling rate. Every participant performed four iterations each of two tasks-- cued picture-naming and code-switching. Utterances in the two conditions (unilingual and mixed) were randomly interspersed within each iteration. In the unilingual condition, the target word was preceded by another English word in the utterance. In the mixed condition, the target was preceded by a Bengali word and thus produced as a switch from Bengali. Every target word appeared twice during a test block; once in a unilingual and once in a mixed condition. Participants alternated between the two tasks, and were allowed to take a short break after each test block.\\

The tasks are described below:

\begin{enumerate}[]
	\item Cued picture-naming: Participants were presented with slides in the following sequence-
	\begin{itemize}
		\item Language cue- a word in either English or Bengali orthography
		\item A picture displayed for 50ms; named by the participant in either English or Bengali, depending on the language cue 
		\item Target -- an English word; read out by the participant
		\item Distracter math problem
	\end{itemize}
	Since all the target words were in English, the language cue was used to manipulate the \textit{preceding} word, giving target words in non-switched (unilingual) and switched (mixed) utterances. The orthographies of English and Bengali are visually distinct, and were therefore used to cue language.

	\item Code-switching: Each slide contained one target word embedded in either an English (unilingual) or a Bengali (mixed; target produced as a switch from Bengali) carrier sentence which was read out by the participant. The carrier sentences are given below:
	\begin{itemize}
		\item Unilingual sentence: That is a yellow [\textit{target word}].
		\item Mixed-language sentence: \textipa{o-\:ta \ae k-\:ta kalo} [\textit{target word}]. 
	\end{itemize}

Gloss and translation for Bengali sentence:

	o-\textipa{\:t}a \quad \textipa{\ae}k-\textipa{\:t}a \quad kalo [\textit{target word}]\\
	that-\textsc{clf}  one-\textsc{clf}  black [\textit{target word}]\\
	`that is a black [\textit{target word}]'\\

	The sound preceding the target word was uniform across all sentences (the final sound in \textit{yellow} is produced as [o] in Indian English), to minimize any possible differences due to coarticulatory effects on the target vowel. A specifier-noun construction was used to ensure identical word order in both languages, giving a uniform prosodic context for the target word across condition. In presenting the mixed-language utterances, we used mixed orthography. \\

\end{enumerate}
\subsection{Acoustic analysis}\label{analysis}
The first iteration of each task was treated as a trial, and excluded from further analyses. An additional 67 tokens (2.79\%) were excluded due to errors in presentation/production, giving 2333 target word tokens for analysis. Recordings were manually segmented and annotated in Praat \citep{boersma2016praat}, and the first and second formant frequencies (F1 and F2) of the target vowels \nt{2, ae} were measured at five points-- in 10\% increments starting at 5\% into the vowel. This gave measures of F1 and F2 at 5\%, 15\%, 25\%, 35\% and 45\% into the vowel, allowing us to observe any dynamic changes in the extent of transfer over the course of producing the vowel.

\section{Results} \label{results}

%Figures ... show the findings of the experiment. 
All statistical analyses were performed in R \citep{r}. The dependent variables were vowel height (F1) and backness (F2). We fitted a series of linear mixed effects models using the \emph{lme4} package \citep{lme4}, to examine the effect of experimental conditions on lobanov normalized F1 and F2 at each of the five points in the vowel (sec.\ref{analysis}). For all mixed models, the alpha criterion was set at $|t| > 2 $. The independent variables or fixed factors, along with their levels, are given in table \ref{table variables}. Subject and Item were treated as random factors.

\begin{table}[h]
	\caption{Fixed factors}
	\sffamily
	\centering
	\begin{tabular}{p{0.5\textwidth}p{0.5\textwidth}}
		\hline 
		Factor & Levels\\ 
		\hline
		Gender & male, female\\
		Vowel & \nt{2}, \nt{\ae} \\
		Task (corresponding to the two paradigms) & p (Picture: cued picture-naming, s (Sentence: code-switching)\\
		Context (corresponding to the experimental conditions) & e (English: unilingual), b (Bengali: mixed)\\ [1ex]
		\hline	
	\end{tabular}
	\label{table variables}
\end{table}

Our target in this analysis was to model vowel quality at different points in the vowel using the same mixed-effects model, allowing us to track if/how transfer changes over time. To do this, we arbitrarily chose a point in the vowel: F1 at 25\% (henceforth F1$\_$25) to identify an optimal model, and then fitted this to all other points.

Vowel quality and gender are expected to systematically affect F1. Thus, we began with a model that had Vowel and Gender as fixed effects, random intercepts for Subject and Item, and random slopes for each within-unit variable that was relevant to our hypotheses-- by-subject slopes for Task, Context, and Vowel; by-item slopes for Task and Context. Next, adding Context as a fixed effect significantly improved the model fit ($\chi^2(1) = 5.54, p= 0.01854$), suggesting that the experimental condition affects vowel height. Following from our hypothesis \textbf{Asymmetry}, we expected the extent of change in vowel height between the conditions to be mediated by the identity of the vowel. Thus, we added a two-way interaction factor between Context and Vowel (Context*Vowel). As expected, this was found to significantly improve model fit ($\chi^2(1) = 12.61, p= 0.0003837$). To assess whether the experimental paradigm affected F1, we added Task as a fixed effect. This marginally improved the model ($\chi^2(1) = 3.74, p= 0.05302$). Towards addressing our research question \textbf{Paradigm}, we wanted to assess whether the extent of shift differs across the two paradigms. However, no significant improvement was achieved by adding an interaction effect between Task and Context. Subsequent additions of other interactions, and removing each of the fixed effects in turn, did not lead to any further improvement.

Context and Vowel are both within-subject factors. Since our hypothesis concerns their interaction, we modified the random effects structure by adding a by-subject random slope for the interaction term \citep{brauer2018linear}. This improved the model fit ($\chi^2(5) = 11.43, p = 0.04342$). Thus, we report on a model that includes Gender, Task, Context, Vowel, Context*Vowel as fixed effects, random intercepts for Subject and Item, by-subject random slopes for Task, Context, Vowel and Context*Vowel, and by-item random slopes for Task and Context.\\

The following sections discuss the results for vowel height and backness respectively:

\subsection{Vowel Height -- F1}

Table \ref{table_f1_fixed_effects} summarizes the results of the fixed effects from the model for F1 at each point. As expected, we observe a significant effect of gender, such that male speakers have a lower mean F1 than female speakers. Vowel is not a significant predictor of F1. This is surprising, given that the phonological category is expected to systematically affect vowel quality. However, a boxplot visualization (Figure\ref{boxplot_F1}) shows that the participants in this study vary greatly both in the relative heights of the two categories \nt{2, \ae} and the extent of height difference between them. Note that these measures are only from the English (unilingual) condition, suggesting that this variability is a feature of the participants' IE vowel categories, independent of the effects of switching. The lack of any independent effect of Context on F1 suggests that using English words in mixed utterances does not lead to a uniform lowering/raising across vowels. Instead, the significant interaction between Context and Vowel indicates a lowering in \nt{2}, but not in \nt{\ae}, in the mixed-language condition. The results also reveal a significant effect of Task, such that the sentence task (code-switching) leads to higher F1 values in both vowels, relative to the cued picture-naming task. This effect is strongest at the beginning of the vowel, and disappears by 45\%. This is an unexpected result, since we have no principled reason to expect code-switching to induce an across-board lowering of vowels. As noted earlier, there is no significant interaction of Task with any of the other fixed factors. We used a series of ANOVAs to compare the model elaborated above with a null model that differs in the absence of the main factor which concerns our hypothesis, Vowel*Context, at each point in the vowel. We find that the difference is not significant at 5\%, and increases over the course of articulating the vowel. 
%ANOVA table add

\begin{figure}[h] \centering

	\includegraphics[scale=0.8]{vowel_by_subject_ggplot}
	\caption{Subject-wise mean F1; unilingual English}
	\label{boxplot_F1}
\end{figure}

\begin{table}[htbp] \centering 
	\caption{Model outputs for F1 at each of the five points in the vowel} 
	\label{table_f1_fixed_effects} 
	\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
		\\[-1.8ex]\hline 
		\hline \\[-1.8ex] 
		& \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
		\cline{2-6} 
		\\[-1.8ex] & F1\_5 & F1\_15 & F1\_25 & F1\_35 & F1\_45 \\ 
		\\[-1.8ex] & (Model 1) & (Model 2) & (Model 3) & (Model 4) & (Model 5)\\ 
		\hline \\[-1.8ex] 
		(Intercept) & 0.447$^{**}$ & 0.718$^{***}$ & 0.750$^{***}$ & 0.863$^{***}$ & 0.921$^{***}$ \\ 
		& (0.211) & (0.176) & (0.153) & (0.152) & (0.150) \\ 
		& & & & & \\ 
		Gender(male) & $-$1.171$^{***}$ & $-$1.526$^{***}$ & $-$1.586$^{***}$ & $-$1.668$^{***}$ & $-$1.681$^{***}$ \\ 
		& (0.237) & (0.184) & (0.168) & (0.177) & (0.175) \\ 
		& & & & & \\ 
		Context(e) & $-$0.010 & $-$0.015 & 0.033 & 0.025 & 0.037 \\ 
		& (0.058) & (0.057) & (0.042) & (0.040) & (0.035) \\ 
		& & & & & \\ 
		Vowel(\textipa{2}) & 0.022 & $-$0.070 & $-$0.034 & $-$0.131 & $-$0.216 \\ 
		& (0.170) & (0.175) & (0.170) & (0.174) & (0.168) \\ 
		& & & & & \\ 
		Task(s) & 0.319$^{***}$ & 0.235$^{***}$ & 0.172$^{**}$ & 0.133$^{*}$ & 0.102 \\ 
		& (0.091) & (0.083) & (0.081) & (0.077) & (0.073) \\ 
		& & & & & \\ 
		Context(e):Vowel(\textipa{2}) & $-$0.126$^{**}$ & $-$0.138$^{**}$ & $-$0.185$^{***}$ & $-$0.191$^{***}$ & $-$0.192$^{***}$ \\ 
		& (0.056) & (0.060) & (0.055) & (0.064) & (0.066) \\ 
		& & & & & \\ 
		\hline \\[-1.8ex] 
		Observations & 2,333 & 2,333 & 2,333 & 2,333 & 2,333 \\ 
		Log Likelihood & $-$1,771.558 & $-$1,589.316 & $-$1,278.310 & $-$1,313.795 & $-$1,380.083 \\ 
		Akaike Inf. Crit. & 3,599.116 & 3,234.631 & 2,612.621 & 2,683.589 & 2,816.165 \\ 
		Bayesian Inf. Crit. & 3,760.253 & 3,395.769 & 2,773.758 & 2,844.727 & 2,977.303 \\ 
		\hline 
		\hline \\[-1.8ex] 
		\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
	\end{tabular} 
\end{table}

\subsection{Vowel backness -- F2}

Table \ref{table_F2_fixed_effects} summarizes the results of fixed effects from the model for F2 at each point. Once again, we see that male speakers have a lower mean F2 than female speakers, as expected. The estimates for Vowel show that excepting the beginning of the vowel, the category \nt{2} is consistently produced as more backed than \nt{\ae}. Unlike vowel height, Context has an independent effect on vowel backness, such that both the vowels in this study have higher F2 values in the Bengali (mixed) condition, i.e. they are fronted in the mixed condition. A significant interaction between Vowel and Context reveals an asymmetry in the extent of fronting between the two vowels -- the shift is greater in the case of \nt{\ae}. We see a significant effect of Task on F2, such that vowels in the Sentence task (code-switching) were more backed than those in the picture-naming task. As with F1, this is an unexpected result, as we have no principled reason to expect code-switching to induce an across-board backing of vowels. A lack of significant interaction with Context suggests that there is no difference in the extent of shift between the two paradigms. The results of ANOVAs comparing the full model to a null model that lacks the Vowel*Context interaction factor reveal that as with vowel height, the change in vowel backness between the two conditions is not significant at 5\%, and increases over the course of the vowel.

\begin{table}[htbp] \centering 
	\caption{Model outputs for F2 at each of the five points in the vowel} 
	\label{table_F2_fixed_effects} 
	\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
		\\[-1.8ex]\hline 
		\hline \\[-1.8ex] 
		& \multicolumn{5}{c}{\textit{Dependent variable:}} \\ 
		\cline{2-6} 
		\\[-1.8ex] & F2\_5 & F2\_15 & F2\_25 & F2\_35 & F2\_45 \\ 
		\\[-1.8ex] & (Model 1) & (Model 2) & (Model 3) & (Model 4) & (Model 5)\\ 
		\hline \\[-1.8ex] 
		(Intercept) & 0.607$^{***}$ & 0.838$^{***}$ & 0.778$^{***}$ & 0.827$^{***}$ & 0.826$^{***}$ \\ 
		& (0.222) & (0.213) & (0.211) & (0.203) & (0.204) \\ 
		& & & & & \\ 
		Gender(male) & $-$0.839$^{***}$ & $-$0.891$^{***}$ & $-$0.807$^{***}$ & $-$0.750$^{***}$ & $-$0.748$^{***}$ \\ 
		& (0.124) & (0.106) & (0.107) & (0.116) & (0.109) \\ 
		& & & & & \\ 
		Context(e) & 0.003 & $-$0.088$^{**}$ & $-$0.065$^{**}$ & $-$0.063 & $-$0.103$^{***}$ \\ 
		& (0.036) & (0.036) & (0.032) & (0.044) & (0.033) \\ 
		& & & & & \\ 
		Vowel(\textipa{2}) & $-$0.217 & $-$0.615$^{***}$ & $-$0.651$^{***}$ & $-$0.801$^{***}$ & $-$0.836$^{***}$ \\ 
		& (0.175) & (0.183) & (0.175) & (0.168) & (0.166) \\ 
		& & & & & \\
		Task(s) & $-$0.234$^{***}$ & $-$0.194$^{***}$ & $-$0.152$^{***}$ & $-$0.141$^{***}$ & $-$0.092$^{**}$ \\ 
		& (0.029) & (0.035) & (0.050) & (0.047) & (0.045) \\ 
		& & & & & \\
		Context(e):Vowel(\textipa{2}) & 0.031 & 0.110$^{***}$ & 0.128$^{***}$ & 0.091$^{**}$ & 0.147$^{***}$ \\ 
		& (0.050) & (0.041) & (0.040) & (0.046) & (0.043) \\ 
		& & & & & \\   
		\hline \\[-1.8ex] 
		Observations & 2,333 & 2,333 & 2,333 & 2,333 & 2,333 \\ 
		Log Likelihood & $-$1,283.531 & $-$1,478.494 & $-$1,174.087 & $-$1,222.682 & $-$1,019.033 \\ 
		Akaike Inf. Crit. & 2,623.062 & 3,012.988 & 2,404.175 & 2,501.363 & 2,094.066 \\ 
		Bayesian Inf. Crit. & 2,784.199 & 3,174.125 & 2,565.312 & 2,662.501 & 2,255.204 \\ 
		\hline 
		\hline \\[-1.8ex] 
		\textit{Note:}  & \multicolumn{5}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
	\end{tabular} 
\end{table}

\section{Discussion}

With regard to our first research question \textbf{L2 vowels}, the results confirm that the L2 vowels of proficient bilinguals can be affected by mixed-language processing. Both vowels in this study showed systematic shifts in the mixed-language condition compared to baseline unilingual productions. In the context of an ICM--based analysis of between-language asymmetries in phonetic transfer as proposed by \cite{olson2013bilingual}, this is a significant finding: it demonstrates that the postulated lower switch cost on an L2 does not entirely mask transfer effects on L2 vowels. Moreover, our findings differ from \cite{tsui2019impact} who report on the basis of VOT measurements that balanced bilinguals do not evidence an effect of language-switching. The authors attribute this to greater experience with language switching, leading to lower switch-costs in both languages. We find that even proficient bilinguals with extensive experience of language-switching demonstrate cross-language transfer effects in vowel production. Thus, following from the discussion in sec. \ref{asymmetries}, it might be fruitful for future studies to focus on vowels as a potential site for transfer. This would both expand the scope of the existing research (cf. sec. \ref{asymmetry between sounds}) and avoid confounding factors arising from independent differences between long- and short-lag VOT languages.\\

We hypothesized an asymmetrical shift between the two vowels (\textbf{Asymmetry}), owing to differences in (i) their positions in the Indian English vowel space, and (ii) the presence/absence of a corresponding category in the L1 Bengali inventory. The significant interaction between Vowel and Context in both F1 and F2 confirm that the experimental condition affected both height and backness differently for each vowel. As expected, we find a greater degree of shift in \nt{2}, possibly owing to its position in a less dense part of the IE vowel space.
Figure \ref{vowels_e_b} shows $F1xF2$ plots for vowel productions in the unilingual and mixed (dotted ellipses) conditions, in comparison to the related Bengali categories (colored ellipses) from \cite{dutta2021}.   
The category \nt{2} is lowered and fronted in the mixed language condition as expected. The direction of shift in \nt{\ae} is surprising-- it is fronted in the mixed-language condition. While phonological constraints on IE suggested that a shift in the F1 dimension could potentially risk the loss of contrast and is thus expected to be minimal, Figure \ref{vowels_e_b} shows that \nt{\ae} in Bengali is not fronted with respect to the English category. Thus, fronting in the mixed condition cannot be explained in relation the former, at least from the present data. Recall that the Bengali vowel productions used here are from a different set of speakers than our participants. Assuming some amount of individual differences in the precise phonetic realization of vowel categories in both languages, this could be a potential source of explanation for this result. Data from a larger number of speakers for both languages, as well as within-speaker comparisons of English and Bengali vowels, are required to clarify this. \\

\begin{figure}[h] \centering
	\includegraphics[scale=0.8]{vowels_e_b_final}
	\caption{Vowel categories in unilinual English (solid lines), mixed condition (dotted lines), and baseline Bengali (colored ellipses)}
	\label{vowels_e_b}
\end{figure}

Our third research question \textbf{Paradigm} concerned the effect of experimental paradigm on the phonetic outcome of mixed-language use. \cite{olson2013bilingual} proposes that in both scripted and spontaneous code-switching paradigms, ``discursive properties may partially drive the production of code-switches, masking underlying effects of interaction", and thus proposes cued picture-naming as a preferable alternative. If such discursive properties independently affect the phonetic outcome of language mixing in an experimental setup, this calls to question the comparability of results across studies using these different paradigms. We verified this by comparing productions of the same items from the same group of participants in a scripted CS paradigm and a cued picture-naming paradigm. We found no significant interaction of Task either with Context, or with the Vowel*Context interaction, suggesting that the extent of phonetic transfer was not affected by the paradigm. 
There are at least two plausible reasons for this-- (i) it is possible that the planning effects of CS do not translate to an increase/decrease in observable transfer at all. This would mean that the pre-switch effects observed, for example, in \cite{muldner2019phonetics} and \cite{bullock2009trying} are better interpreted as resulting from some other aspect of the experimental setup, than the act of code-switching itself. While planning could additionally affect the words preceding the switch, the effects on the switched token itself do not change. (ii) Our study used a frame sentence to induce code switching, and not all the items led to pragmatically meaningful sentences. Thus, another possible explanation is that the discursive effects discussed by \cite{olson2013bilingual} are triggered by more naturalistic paradigms where attention to meaning and interlocutor are important. More within-study comparisons with less structured paradigms, and spontaneous code-switching data, can clarify this. The results of the present study indicate that for the purposes of observing dynamic phonetic transfer in an experimental setup, elicitation through cued-picture-naming vs (controlled) code-switching does not significantly alter the behavior under study. Thus, the choice of paradigm can be based on other considerations. Moreover, this suggests that (all else being equal) the results from such studies can be compared directly.

The paradigm did affect vowel quality, but this applied uniformly across experimental conditions: all vowels were more backed and lowered in the code-switching task. Given that at least for F1, the effect is limited to the early stages of articulating the vowel, it is plausible that this lowering is an artifact of the particular stimuli used here---resulting from some corarticulatory effect that was absent in the picture-naming task. While this is intriguing, it is beyond the scope of the present study.  \\

\section{Conclusion}
This study examined the phonetic outcome of mixed-language production in a group of highly proficient bilingual speakers of Bengali and Indian English. Specifically, we examined the spectral quality of two L2 vowel categories \nt{2} and \nt{\ae} produced in two experimental conditions: unilingual (English) and mixed (switching from Bengali to English).
Comparisons of F1 and F2 reveal that mixed language processing temporarily increases phonetic interaction between the languages, causing \nt{2} to converge towards a related L1 category \nt{a:}. The extent and direction of such shift is asymmetrical across the two categories, which parallels findings in prior studies. This asymmetry appears to be mediated by the language-specific distribution of vowel categories. Mixed productions were elicited in two switching paradigms: code-switching and cued picture-naming, to assess whether the postulated discursive elements in code-switching independently influences the outcome of phonetic transfer. Results suggest that in an experimental setup, elicitation through cued-picture-naming vs (controlled) code-switching does not significantly alter the behavior under study. \\

This study has a number of limitations: (i) the `baseline' values for Bengali vowel categories were estimated from a different set of speakers than the participants of this study, potentially obscuring the role of individual differences in the realization of vowel categories; (ii) we used a frame sentence to elicit code-switched productions such that not all the sentences were pragmatically well-formed. This could have blocked some of the discursive effects that might result from speech planning in ordinary connected speech, thus obscuring differences between the two paradigms under study. Future work should address this by considering data from more naturalistic code-switching paradigms.

\section*{Abbreviations}
\textsc{clf} = classifier

\section*{Data availability}
The data files and analysis code are available at:  \url{https://osf.io/dsb2x/?view_only=5cb821070bc74a038a293239157c3b11}

\section*{Competing interests}
The authors have no competing interests to declare.

%\printbibliography %for use with biblatex; comment out if you use natbib
\bibliography{references} %for use with natbib; comment out if you use biblatex, and change 'sample' by the name of your bib-file

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% TeX-engine: luatex
%%% End: