### 2022-02-20 RCode/transfer_analysis.Rmd
analysis log-mean normalization, rmd file

---
title: "Phonetic Tansfer analysis code"
author: "Auromita"
date: "1/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
```

Load packages and data, create objects for lobanov normalized F1 and F2, create dataframe.

(note: file paths relative to project root folder (phonetic_transfer))

```{r, include=FALSE}

library(tidyverse)
library(phonR)
library(lme4)
library(lmerTest)
library(stargazer)
library(pwr)
library(smir)

df <- read.csv("RCode/formant_data_5.csv", check.names = TRUE) %>%
  rename_all(tolower) %>%
  mutate_if(is.character,as.factor)

str(df)

attach(df)

# pool measures from all 5 points in the vowel, then from that-- calculate the average F1 for each speaker, and normalize all data points from that speaker wrt that.

normF1 <- with(df,normLobanov(cbind(f1_5,f1_15,f1_25,f1_35,f1_45)),group=subject)
normF2 <- with(df,normLobanov(cbind(f2_5,f2_15,f2_25,f2_35,f2_45)),group=subject)

df_lobanov <- data.frame(subject,gender,task,word,context,vowel,normF1,normF2)

detach(df)

rm(normF1,normF2)

# data frame with time as variable
df_lobanov_time <- df_lobanov %>%
  rename(f1_1=f1_5, f1_2=f1_15, f1_3=f1_25, f1_4=f1_35, f1_5=f1_45, f2_1=f2_5, f2_2=f2_15, f2_3=f2_25, f2_4=f2_35, f2_5=f2_45) %>%
  # dplyr pivot() functions: convert data between long(indexed) and wide(cartesian) formats
  pivot_longer(starts_with("f"),
               names_to = c(".value", "time"),
               names_sep = "_",
               names_transform = list(time = as.factor)
               )
# Since there are multiple observations per row (F1 and F2) that need to go in separate columns in the output, the ".value" lets you specify that part of the variable name will be used to split the data in the output. The rest of the name becomes the variable "time"

rm(df_lobanov)
```

Log-mean normalization:

```{r}

# data in long format for log-mean normalization 
df_long <- df %>%
  select(!c(vowel_duration, f1_mid)) %>%
  rename(f1_1=f1_5, f1_2=f1_15, f1_3=f1_25, f1_4=f1_35, f1_5=f1_45, f2_1=f2_5, f2_2=f2_15, f2_3=f2_25, f2_4=f2_35, f2_5=f2_45, iteration=filename) %>%
  pivot_longer(starts_with("f"), # pivot all the columns whose names start with "f"
               values_to = "ff_raw",
               # drop the f part, store the rest as values for new variables "formant" and "time"
               names_to = c(NA, "formant", "time"), 
               # how to divide up the variable name (f + formant + _ + time)
               names_pattern = "(.)(.)_(.)") %>% 
  # reorder columns
  relocate("subject", "gender", "task", "word", "context", "vowel", "time", "formant", "ff_raw") %>%
  mutate_if(is.character, as.factor) %>%
  # new column with log-transformed FFs
  mutate(ff_log = log(ff_raw)) %>%
  # composite variable vowel-formant (estimate for the dialectal reference pattern-- the target ff for that vowel in the dialect)
  mutate (vowel.formant = factor(interaction(vowel, formant)))

# fitting a regression model to predict log-transformed formant-frequencies (ff-log) for a given vowel, formant and speaker as the sum of a speaker displacement term (subject) and normalized vowel-formant effects (vowel.formant)  (Barreda&Neary 2018)

### only vowel midpoints for figures
df_long <- df_long %>%
  filter(time==5)

attach(df_long)

# forcing the intercept to 0 so that coefficients are calculated for all levels of "subject"
M = lm (ff_log ~ 0 + subject + vowel.formant, contrasts = list(vowel.formant = contr.sum))
summary(M)

# coefficients of the speaker effect (estimate of speaker parameter)
speaker_coeff = dummy.coef(M)$subject
# coefficients of the vowel.formant effect (estimate of dialect reference/target formant for each vowel)
vowel_coeff = dummy.coef(M)$vowel.formant

detach(df_long)

# from each log-transformed ff of a particular speaker, want to subtract the speaker coefficient

temp <- as.data.frame(speaker_coeff) %>%
  rownames_to_column(var = "subject")

df_long <- df_long %>%
  full_join(temp, by = "subject")

df_long <- df_long %>%
  mutate(ff_log_normalized = ff_log - speaker_coeff)

df_long <- df_long %>%
  mutate(ff_exp_normalized = exp(ff_log_normalized))

rm(temp,speaker_coeff,vowel_coeff)

# putting F1 and F2 in separate columns for plotting (pivot_wider doesn't work because there are no unique identifiers for each vowel, since there were multiple reps in each condition). So crude workaround:

df_clean <- df_long %>%
  select(subject, gender, task, iteration, word, context, vowel, time, formant, ff_log_normalized, ff_exp_normalized, ff_raw)

f1 <- df_clean %>%
  filter(formant == 1)

f2 <- df_clean %>%
  filter(formant == 2)

df_normalized <- bind_cols(f1,f2$ff_log_normalized,f2$ff_exp_normalized, f2$ff_raw)%>%
  mutate(formant=NULL) %>%
  rename(f1_log = ff_log_normalized, f1_exp = ff_exp_normalized, f1_raw = ff_raw, f2_log =...13, f2_exp = ...14, f2_raw = ...15)

rm(df_clean, f1, f2)

```

Baseline Bengali data from the Shruti corpus:

```{r, include=FALSE}

# list of files in directory (paths relative to project root)
file_list <- list.files(path="shruti/merged_speaker_data", full.names = T)
vowel_list <- "oh|u|i|ee|A|E|aa"

# create dataframe by reading data from each file in folder
df_ben <- file_list %>%
  set_names(.) %>%
  # apply read_table() to each file, read only specified columns
  map_df(read_table, .id = "FileName", col_types = cols_only(rowLabel ='f', Time ='d', F1 ='d', F2 ='d')) %>%
  # rename columns
  rename(speaker=FileName, vowel=rowLabel, time_absolute=Time, f1=F1, f2=F2) %>%
  # only keep rows that have vowels
  filter(grepl(vowel_list, vowel)) %>%
  # clean column name (remove file path)
  mutate(speaker = str_replace(speaker, "shruti/merged_speaker_data/", "")) %>%
  mutate(speaker = str_replace(speaker, "_merged.txt", "")) %>%
  mutate(speaker = as.factor(speaker)) %>%
  droplevels()

# columns for gender and time point (1-20)

male = c("abd", "chandan", "kunal", "plb", "sayan", "sudha", "swarnendu", "amitava", "deb", "mainak", "rajashree", "sm", "suman", "xxx", "bd", "jit", "moy", "samaresh", "smt", "suparna", "chan", "jyotirmoy", "padma", "sandi", "styabrata", "suparnakdas")

female = c("manika", "msm", "punam", "rita",  "ritwika",  "shyamoshree",  "suranjana",  "tml")

timepoints <- c(1:20)

# add column for gender using mutate() and ifelse()
# add column for time (iterate over 1:20; length.out: desired length of final vector (here- total no. of rows))
df_ben <- df_ben %>%
  mutate(gender = ifelse(speaker %in% male, "male", "female"),
         time = rep(timepoints, length.out = 2458340), .after = time_absolute,
         time = as.factor(time)
         )

# Lobanov normalization (normalizing over all vowels here. For the English data-- only have 2 vowels.)

# attach(bengali_df)
# lobanov <- with(bengali_df,normLobanov(cbind(f1,f2), group = speaker))
# bengali_lobanov <- data.frame(speaker, gender, vowel, time, lobanov)
# rm(lobanov)
# detach(bengali_df)

rm(female, file_list, male, timepoints, vowel_list)

```

Log-mean normalization for Bengali data:

```{r}

# subset of dataframe with only vowels of interest (A and E) at vowel midpoints (timepoints 8-12),
# removing outlier formant values (<300 Hz): ~ 11% of rows. In full dataset (all vowels): ~14%

df_ben_AE <- df_ben %>%
  filter(grepl("A|E", vowel), time %in% (8:12), f1>500) %>%
  droplevels() %>%
  mutate(time_absolute = NULL) %>%
  relocate(gender, .after = speaker)

# data in long format for log-mean normalization

df_ben_long <- df_ben_AE %>%
  pivot_longer(starts_with("f"), # pivot all the columns whose names start with "f"
               values_to = "ff_raw",
               # drop the f part, store the rest as value for new variable "formant" 
               names_to = c(NA, "formant"), 
               # how to divide up the variable name (f + formant)
               names_pattern = "(.)(.)") %>% 
  mutate_if(is.character, as.factor) %>%
  # new column with log-transformed FFs
  mutate(ff_log = log(ff_raw)) %>%
  # composite variable vowel-formant (estimate for the dialectal reference pattern-- the target ff for that vowel in the dialect)
  mutate (vowel.formant = factor(interaction(vowel, formant)))

# fitting a regression model to predict log-transformed formant-frequencies (ff-log) for a given vowel, formant and speaker as the sum of a speaker displacement term (subject) and normalized vowel-formant effects (vowel.formant)  (Barreda&Neary 2018)

attach(df_ben_long)

# forcing the intercept to 0 so that coefficients are calculated for all levels of "subject"
M_ben = lm (ff_log ~ 0 + speaker + vowel.formant, contrasts = list(vowel.formant = contr.sum))
summary(M)

# coefficients of the speaker effect (estimate of speaker parameter)
speaker_coeff = dummy.coef(M_ben)$speaker
# coefficients of the vowel.formant effect (estimate of dialect reference/target formant for each vowel)
vowel_coeff = dummy.coef(M_ben)$vowel.formant

detach(df_ben_long)

# from each log-transformed ff of a particular speaker, want to subtract the speaker coefficient

temp <- as.data.frame(speaker_coeff) %>%
  rownames_to_column(var = "speaker")

df_ben_long <- df_ben_long %>%
  full_join(temp, by = "speaker")

df_ben_long <- df_ben_long %>%
  mutate(ff_log_normalized = ff_log - speaker_coeff)

df_ben_long <- df_ben_long %>%
  mutate(ff_exp_normalized = exp(ff_log_normalized))

rm(temp,speaker_coeff,vowel_coeff)

# putting F1 and F2 in separate columns for plotting (pivot_wider doesn't work because there are no unique identifiers for each vowel, since there were multiple reps in each condition). So crude workaround:

df_clean <- df_ben_long %>%
  select(speaker, gender, vowel, time, formant, ff_log_normalized, ff_exp_normalized)

f1 <- df_clean %>%
  filter(formant == 1)

f2 <- df_clean %>%
  filter(formant == 2)

df_ben_normalized <- bind_cols(f1, f2$ff_log_normalized, f2$ff_exp_normalized) %>%
  mutate(formant=NULL) %>%
  rename(f1_log=ff_log_normalized, f1_exp=ff_exp_normalized, f2_log=...8, f2_exp=...9)

rm(df_clean, f1, f2)

variance <- var(df_ben$f1)
variancef2 <- var(df_ben$f2)

```

Plotting vowel data:

```{r}

# IPA vowel symbols
remapping_english <- c(`\\ae` = "æ", `\\vt` = "ʌ")
remapping_bengali <- c(A = "a:", E = "æ")

# add column with unicode vowels
df_ben_normalized$unicodevowel <- remapping_bengali[as.character(df_ben_normalized$vowel)]

df_normalized$unicodevowel <- remapping_english[as.character(df_normalized$vowel)]

# plot: unilingual English, mixed, and baseline Bengali
xlim <- c(1,-0.2)
ylim = c(0,-1)

# Bengali vowels
with(df_ben_normalized,plotVowels(f1_log, f2_log, vowel,
                        plot.tokens = F, plot.means = TRUE, pch.tokens = unicodevowel, pch.means = unicodevowel, 
                        var.col.by = vowel, var.sty.by = vowel, 
                        cex.tokens = 0.8, cex.means = 1, alpha.means = 0.8,alpha.tokens = 0.3,
                        ellipse.line = F, ellipse.fill = T, fill.opacity = 0.5,
                        poly.line = F, poly.order = c("A","E"),
                        xlim = xlim, ylim = ylim, 
                        legend.kwd="bottomleft", legend.args = list(legend=c("baseline æ","baseline a:")), 
                        axes=FALSE, pretty = TRUE))
par(new=TRUE)

# English vowels
with(df_normalized, plotVowels(f1_log, f2_log, vowel, group = context, 
                               plot.tokens = F, plot.means = TRUE , pch.tokens = unicodevowel, pch.means = unicodevowel,
                               var.col.by = NA, var.sty.by = context, 
                               cex.tokens = 0.8, cex.means = 1, alpha.means = 0.5,alpha.tokens = 0.3,
                               ellipse.line = T,  ellipse.fill = T,
                               poly.line = F, poly.order = c("\\ae","\\vt"),
                               xlim = xlim, ylim = ylim, 
                               legend.kwd = "bottomright", legend.args = list(legend=c("unilingual","mixed")),
                               fill.opacity = 0.1, pretty = TRUE))

```

LME models:

```{r}

## main model

phon_acco.model = lmer(f1 ~ 
                         gender + context*vowel + context*task +
                         (1|word) + (1+context*vowel|subject), 
                         data= df_time, REML=FALSE)
summary(phon_acco.model)
summary(rePCA(phon_acco.model))

fixef(phon_acco.model)
powerSim(phon_acco.model)

trial <- lmer(f1 ~
                context*vowel + gender + task +
                (1+context*vowel|subject) + (1|word),
                data = df_time, REML = FALSE)

summary(trial)

fixef(trial)["contexte:tasks"] <- 0.1
powerSim(trial, test = fixed("tasks"), nsim = 20)
pc <- powerCurve (trial, along = "subject", test = fixed("task"), nsim = 20)
plot(pc)

null.model = lmer(f1 ~ 
                    gender + context*vowel + task + 
                    (1|word) + (1+context*vowel|subject), 
                    data=df_time, REML=FALSE)
summary(null.model)

anova(phon_acco.model,null.model)

coef(phon_acco.model)

```