### 2022-02-26 RCode/transfer_analysis.Rmd
analysis models

---
title: "Phonetic Tansfer analysis code"
author: "Auromita"
date: "1/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
```

Load packages and data, create objects for lobanov normalized F1 and F2, create dataframe.

(note: file paths relative to project root folder (phonetic_transfer))

```{r, include=FALSE}

library(tidyverse)
library(phonR)
library(lme4)
library(lmerTest)
library(pwr)
library(simr)

df <- read.csv("RCode/formant_data_5.csv", check.names = TRUE) %>%
  rename_all(tolower) %>%
  mutate_if(is.character,as.factor)

str(df)

attach(df)

# pool measures from all 5 points in the vowel, then from that-- calculate the average F1 for each speaker, and normalize all data points from that speaker wrt that.

normF1 <- with(df,normLobanov(cbind(f1_5,f1_15,f1_25,f1_35,f1_45)),group=subject)
normF2 <- with(df,normLobanov(cbind(f2_5,f2_15,f2_25,f2_35,f2_45)),group=subject)

df_lobanov <- data.frame(subject,gender,task,word,context,vowel,normF1,normF2)

detach(df)

rm(normF1,normF2)

# data frame with time as variable

# lobanov-normalized FFs
df_lobanov_time <- df_lobanov %>%
  rename(f1_1=f1_5, f1_2=f1_15, f1_3=f1_25, f1_4=f1_35, f1_5=f1_45, f2_1=f2_5, f2_2=f2_15, f2_3=f2_25, f2_4=f2_35, f2_5=f2_45) %>%
  # dplyr pivot() functions: convert data between long(indexed) and wide(cartesian) formats
  pivot_longer(starts_with("f"),
               names_to = c(".value", "time"),
               names_sep = "_",
               names_transform = list(time = as.factor)
               )
# Since there are multiple observations per row (F1 and F2) that need to go in separate columns in the output, the ".value" lets you specify that part of the variable name will be used to split the data in the output. The rest of the name becomes the variable "time"

# raw FFs
df_time <- df %>%
  select(!c("filename", "f1_mid", "vowel_duration")) %>%
  rename(f1_1=f1_5, f1_2=f1_15, f1_3=f1_25, f1_4=f1_35, f1_5=f1_45, f2_1=f2_5, f2_2=f2_15, f2_3=f2_25, f2_4=f2_35, f2_5=f2_45) %>%
  # dplyr pivot() functions: convert data between long(indexed) and wide(cartesian) formats
  pivot_longer(starts_with("f"),
               names_to = c(".value", "time"),
               names_sep = "_",
               names_transform = list(time = as.factor)
               )

```

Log-mean normalization:

```{r, eval=FALSE, include=FALSE}

# data in long format for log-mean normalization 
df_long <- df %>%
  select(!c(vowel_duration, f1_mid)) %>%
  rename(f1_1=f1_5, f1_2=f1_15, f1_3=f1_25, f1_4=f1_35, f1_5=f1_45, f2_1=f2_5, f2_2=f2_15, f2_3=f2_25, f2_4=f2_35, f2_5=f2_45, iteration=filename) %>%
  pivot_longer(starts_with("f"), # pivot all the columns whose names start with "f"
               values_to = "ff_raw",
               # drop the f part, store the rest as values for new variables "formant" and "time"
               names_to = c(NA, "formant", "time"), 
               # how to divide up the variable name (f + formant + _ + time)
               names_pattern = "(.)(.)_(.)") %>% 
  # reorder columns
  relocate("subject", "gender", "task", "word", "context", "vowel", "time", "formant", "ff_raw") %>%
  mutate_if(is.character, as.factor) %>%
  # new column with log-transformed FFs
  mutate(ff_log = log(ff_raw)) %>%
  # composite variable vowel-formant (estimate for the dialectal reference pattern-- the target ff for that vowel in the dialect)
  mutate (vowel.formant = factor(interaction(vowel, formant)))

# fitting a regression model to predict log-transformed formant-frequencies (ff-log) for a given vowel, formant and speaker as the sum of a speaker displacement term (subject) and normalized vowel-formant effects (vowel.formant)  (Barreda&Neary 2018)

### only vowel midpoints for figures
df_long <- df_long %>%
  filter(time==5)

attach(df_long)

# forcing the intercept to 0 so that coefficients are calculated for all levels of "subject"
M = lm (ff_log ~ 0 + subject + vowel.formant, contrasts = list(vowel.formant = contr.sum))
summary(M)

# coefficients of the speaker effect (estimate of speaker parameter)
speaker_coeff = dummy.coef(M)$subject
# coefficients of the vowel.formant effect (estimate of dialect reference/target formant for each vowel)
vowel_coeff = dummy.coef(M)$vowel.formant

detach(df_long)

# from each log-transformed ff of a particular speaker, want to subtract the speaker coefficient

temp <- as.data.frame(speaker_coeff) %>%
  rownames_to_column(var = "subject")

df_long <- df_long %>%
  full_join(temp, by = "subject")

df_long <- df_long %>%
  mutate(ff_log_normalized = ff_log - speaker_coeff)

df_long <- df_long %>%
  mutate(ff_exp_normalized = exp(ff_log_normalized))

rm(temp,speaker_coeff,vowel_coeff)

# putting F1 and F2 in separate columns for plotting (pivot_wider doesn't work because there are no unique identifiers for each vowel, since there were multiple reps in each condition). So crude workaround:

df_clean <- df_long %>%
  select(subject, gender, task, iteration, word, context, vowel, time, formant, ff_log_normalized, ff_exp_normalized, ff_raw)

f1 <- df_clean %>%
  filter(formant == 1)

f2 <- df_clean %>%
  filter(formant == 2)

df_normalized <- bind_cols(f1,f2$ff_log_normalized,f2$ff_exp_normalized, f2$ff_raw)%>%
  mutate(formant=NULL) %>%
  rename(f1_log = ff_log_normalized, f1_exp = ff_exp_normalized, f1_raw = ff_raw, f2_log =...13, f2_exp = ...14, f2_raw = ...15)

rm(df_clean, f1, f2)

```

Baseline Bengali data from the Shruti corpus:

```{r, include=FALSE}

# list of files in directory (paths relative to project root)
file_list <- list.files(path="shruti/merged_speaker_data", full.names = T)
vowel_list <- "oh|u|i|ee|A|E|aa"

# create dataframe by reading data from each file in folder
df_ben <- file_list %>%
  set_names(.) %>%
  # apply read_table() to each file, read only specified columns
  map_df(read_table, .id = "FileName", col_types = cols_only(rowLabel ='f', Time ='d', F1 ='d', F2 ='d')) %>%
  # rename columns
  rename(speaker=FileName, vowel=rowLabel, time_absolute=Time, f1=F1, f2=F2) %>%
  # only keep rows that have vowels
  filter(grepl(vowel_list, vowel)) %>%
  # clean column name (remove file path)
  mutate(speaker = str_replace(speaker, "shruti/merged_speaker_data/", "")) %>%
  mutate(speaker = str_replace(speaker, "_merged.txt", "")) %>%
  mutate(speaker = as.factor(speaker)) %>%
  droplevels()

# columns for gender and time point (1-20)

male = c("abd", "chandan", "kunal", "plb", "sayan", "sudha", "swarnendu", "amitava", "deb", "mainak", "rajashree", "sm", "suman", "xxx", "bd", "jit", "moy", "samaresh", "smt", "suparna", "chan", "jyotirmoy", "padma", "sandi", "styabrata", "suparnakdas")

female = c("manika", "msm", "punam", "rita",  "ritwika",  "shyamoshree",  "suranjana",  "tml")

timepoints <- c(1:20)

# add column for gender using mutate() and ifelse()
# add column for time (iterate over 1:20; length.out: desired length of final vector (here- total no. of rows))
df_ben <- df_ben %>%
  mutate(gender = ifelse(speaker %in% male, "male", "female"),
         time = rep(timepoints, length.out = 2458340), .after = time_absolute,
         time = as.factor(time)
         )

# Lobanov normalization (normalizing over all vowels here. For the English data-- only have 2 vowels.)
# attach(bengali_df)
# lobanov <- with(bengali_df,normLobanov(cbind(f1,f2), group = speaker))
# bengali_lobanov <- data.frame(speaker, gender, vowel, time, lobanov)
# rm(lobanov)
# detach(bengali_df)

rm(female, file_list, male, timepoints, vowel_list)

```

Log-mean normalization for Bengali data:

```{r}

# subset of dataframe with only vowels of interest (A and E) at vowel midpoints (timepoints 8-12),
# removing outlier formant values (<300 Hz): ~ 11% of rows. In full dataset (all vowels): ~14%

df_ben_AE <- df_ben %>%
  filter(grepl("A|E", vowel), time %in% (8:12), f1>600) %>%
  droplevels() %>%
  mutate(time_absolute = NULL) %>%
  relocate(gender, .after = speaker)

# data in long format for log-mean normalization

df_ben_long <- df_ben_AE %>%
  pivot_longer(starts_with("f"), # pivot all the columns whose names start with "f"
               values_to = "ff_raw",
               # drop the f part, store the rest as value for new variable "formant" 
               names_to = c(NA, "formant"), 
               # how to divide up the variable name (f + formant)
               names_pattern = "(.)(.)") %>% 
  mutate_if(is.character, as.factor) %>%
  # new column with log-transformed FFs
  mutate(ff_log = log(ff_raw)) %>%
  # composite variable vowel-formant (estimate for the dialectal reference pattern-- the target ff for that vowel in the dialect)
  mutate (vowel.formant = factor(interaction(vowel, formant)))

# fitting a regression model to predict log-transformed formant-frequencies (ff-log) for a given vowel, formant and speaker as the sum of a speaker displacement term (subject) and normalized vowel-formant effects (vowel.formant)  (Barreda&Neary 2018)

attach(df_ben_long)

# forcing the intercept to 0 so that coefficients are calculated for all levels of "subject"
M_ben = lm (ff_log ~ 0 + speaker + vowel.formant, contrasts = list(vowel.formant = contr.sum))
summary(M)

# coefficients of the speaker effect (estimate of speaker parameter)
speaker_coeff = dummy.coef(M_ben)$speaker
# coefficients of the vowel.formant effect (estimate of dialect reference/target formant for each vowel)
vowel_coeff = dummy.coef(M_ben)$vowel.formant

detach(df_ben_long)

# from each log-transformed ff of a particular speaker, want to subtract the speaker coefficient

temp <- as.data.frame(speaker_coeff) %>%
  rownames_to_column(var = "speaker")

df_ben_long <- df_ben_long %>%
  full_join(temp, by = "speaker")

df_ben_long <- df_ben_long %>%
  mutate(ff_log_normalized = ff_log - speaker_coeff)

df_ben_long <- df_ben_long %>%
  mutate(ff_exp_normalized = exp(ff_log_normalized))

rm(temp,speaker_coeff,vowel_coeff)

# putting F1 and F2 in separate columns for plotting (pivot_wider doesn't work because there are no unique identifiers for each vowel, since there were multiple reps in each condition). So crude workaround:

df_clean <- df_ben_long %>%
  select(speaker, gender, vowel, time, formant, ff_log_normalized, ff_exp_normalized, ff_raw)

f1 <- df_clean %>%
  filter(formant == 1)

f2 <- df_clean %>%
  filter(formant == 2)

df_ben_normalized <- bind_cols(f1, f2$ff_log_normalized, f2$ff_exp_normalized, f2$ff_raw) %>%
  mutate(formant=NULL) %>%
  rename(f1_log=ff_log_normalized, f1_exp=ff_exp_normalized, f1_raw=ff_raw, f2_log=...9, f2_exp=...10, f2_raw=...11)

rm(df_clean, f1, f2)

```

Plotting vowel data:

```{r}

# IPA vowel symbols
remapping_english <- c(`\\ae` = "æ", `\\vt` = "ʌ")
remapping_bengali <- c(A = "a:", E = "æ")

# add column with unicode vowels
df_ben_normalized$unicodevowel <- remapping_bengali[as.character(df_ben_normalized$vowel)]

df_normalized$unicodevowel <- remapping_english[as.character(df_normalized$vowel)]

# plot: unilingual English, mixed, and baseline Bengali
xlim <- c(2500,600)
ylim = c(1100,400)

temp <- df_ben %>%
  filter(time %in% (8:12), f1>300)

with(temp,plotVowels(f1, f2, vowel,
                        plot.tokens = F, plot.means = TRUE, pch.tokens = vowel, pch.means = vowel, 
                        var.col.by = vowel, var.sty.by = vowel, 
                        cex.tokens = 0.8, cex.means = 1, alpha.means = 0.8,alpha.tokens = 0.3,
                        ellipse.line = F, ellipse.fill = T, fill.opacity = 0.3,
                        poly.line = F, poly.order = c("A","E"),
                        xlim = xlim, ylim = ylim, 
                        legend.kwd="bottomleft", legend.args = list(legend=c("baseline æ","baseline a:")), 
                        axes=T, pretty = TRUE))

# Bengali vowels
with(df_ben_normalized,plotVowels(f1_raw, f2_raw, vowel,
                        plot.tokens = F, plot.means = TRUE, pch.tokens = unicodevowel, pch.means = unicodevowel, 
                        var.col.by = vowel, var.sty.by = vowel, 
                        cex.tokens = 0.8, cex.means = 1, alpha.means = 0.8,alpha.tokens = 0.3,
                        ellipse.line = F, ellipse.fill = T, fill.opacity = 0.2,
                        poly.line = F, poly.order = c("A","E"),
                        xlim = xlim, ylim = ylim, 
                        legend.kwd="bottomleft", legend.args = list(legend=c("baseline æ","baseline a:")), 
                        axes=FALSE, pretty = TRUE))
par(new=TRUE)

# English vowels
with(df_normalized, plotVowels(f1_raw, f2_raw, vowel, group = context, 
                               plot.tokens = F, plot.means = TRUE , pch.tokens = unicodevowel, pch.means = unicodevowel,
                               var.col.by = NA, var.sty.by = context, 
                               cex.tokens = 0.8, cex.means = 1, alpha.means = 0.5,alpha.tokens = 0.3,
                               ellipse.line = T,  ellipse.fill = T,
                               poly.line = F, poly.order = c("\\ae","\\vt"),
                               xlim = xlim, ylim = ylim, 
                               legend.kwd = "bottomright", legend.args = list(legend=c("unilingual","mixed")),
                               fill.opacity = 0.1, pretty = TRUE))

```

## LME models:

```{r}
# plot for English vowels:

# Lobanov-normalized FFs
xlim = c(2.5, -2)
ylim = c(2.3, -2)

with(df_lobanov_time, plotVowels(f1, f2, vowel, group = task,
                               plot.tokens = F, plot.means = TRUE , pch.tokens = vowel, pch.means = vowel,
                               var.col.by = task, var.sty.by = NULL, 
                               cex.tokens = 0.8, cex.means = 1, alpha.means = 0.5,alpha.tokens = 0.3,
                               ellipse.line = F,  ellipse.fill = T,
                               poly.line = F, poly.order = c("\\ae","\\vt"),
                               xlim = xlim, ylim = ylim, 
                               legend.kwd = "bottomright", legend.args = gender,
                               fill.opacity = 0.1, pretty = TRUE))

# raw FFs
with(df_time, plotVowels(f1, f2, vowel, group = task, 
                               plot.tokens = F, plot.means = TRUE , pch.tokens = vowel, pch.means = vowel,
                               var.col.by = task, var.sty.by = context, 
                               cex.tokens = 0.8, cex.means = 1, alpha.means = 0.5,alpha.tokens = 0.3,
                               ellipse.line = T,  ellipse.fill = T,
                               poly.line = F, poly.order = c("\\ae","\\vt"),
                               xlim = c(2500, 700), ylim = c(1400, 100), 
                               legend.kwd = "bottomright", legend.args = gender,
                               fill.opacity = 0.1, pretty = TRUE))

```

Gender difference is not removed after lobanov normalization. Normalizing by gender:

```{r}

df_lobanov_gender_time <- df_time %>%
  mutate(f1 = with(df_lobanov_time, normLobanov(f1), group = gender),
         f2 = with(df_lobanov_time, normLobanov(f2), group = gender))

```

## main model

writeup-- (using data from all 5 points. This should be okay because the pre-vocalic consonants are all the same, so any effect of the consonant should be uniform. A difference between contexts in the early part of the vowel suggests differences in target FFs)

Gender effects: try:
- normalize by gender
- use log-mean normalized values

Hypotheses:

- FFs are different across vowels
- FFS are different across contexts
- the two vowel categories are differently affected (precise direction/extent of shift in FFs depends on category, position in vowel space, presence/absence of contrast in L1)
So main effect of interest: context*vowel

- if transfer patterns differ across tasks, then expect: context\*task interaction, OR context\*vowel\*task interaction

- FFs should change across time, so effect of time on FF. But what time is doing to the FF (direction of change)-- depends on the vowel category. So expect: time\*vowel interaction

- if transfer patterns change across time, then expect: context\*time interaction, OR context\*vowel\*time interaction

Random effects structure:

- subject: intercept, context, vowel (target FF different across speakers), context\*vowel interaction (transfer pattern different across speakers)

-item: intercept (FF for vowel different in each item), context (the effect of context is different for each word), NOT vowel (between-item variable), NOT gender (even though within-item, no reason to believe that effect of gender on FF is different for each item)

```{r, results='hide'}
temp <- df_lobanov_time %>%
  filter(time == "5")

# F1

# full model
phon_acco.model = lmer(f1 ~ 
                         gender + time*vowel + vowel*context*task +
                         (1|word) + (1+context+vowel|subject), 
                         data= df_lobanov_time, REML=FALSE)
summary(phon_acco.model)
summary(rePCA(phon_acco.model))

# null model
null.model = lmer(f1~
                   gender + time*vowel + vowel*context + task*vowel +
                   (1|word) + (1+context+vowel|subject), 
                   data= df_lobanov_time, REML=FALSE)

summary(null.model)

anova(phon_acco.model, null.model)
```

Random effects structure: (1+context|word) + (1+context+vowel|subject). But by-word random slope for context is explaining barely any variance, and null model is not converging. So removing that.

- Full model1: gender + time\*vowel + vowel\*context
- null model1: gender + time*vowel
full model is better
- null model2: gender + time*vowel + context
full model still better

Shows: F1 differs across contexts (mediated by vowel category). So transfer is happening.

Now want to see: effect of task?
If there is a difference in transfer patterns across tasks, then expect task to interact with the vowel\*context variable 
- full model: gender + time\*vowel + vowel\*context\*task 
-task by itself affects F1 (unexpected)
- marginal effect of context\*task (effect of context on F1 is moderated by task) -- expected
- significant effect of vowel\*context\*task (transfer pattern is moderated by task) -- expected

- null model: gender + time\*vowel + vowel\*context
full model better

- null model: gender + time\*vowel + vowel\*context + task
(because adding task is significant on its own, and adding it as an interaction term increases model parameters, want to see if having it as an interaction significantly improves fit)
full model better -- task does interact with context and vowel

- null model: gender + time\*vowel + vowel\*context + task\*context 
to see if three-way interaction is needed. Is task only interacting with context?
nope, context\*task not significant here
full model better

gender + time\*vowel + vowel\*context + task\*vowel 
to see if three-way interaction is needed. Is task only interacting with vowel?
context\*vowel is significant
but anova shows that the full model is still better

So: need the task\*context\*vowel interaction

Want to see: is the effect of task local (decreases with time)? Since the effect of juat task on F1 is unexpected, want to see if it might be an artifact of the paradigm/stimuli

```{r}
# F2
phon_acco.model = lmer(f1 ~ 
                         gender + context*vowel + context*task +
                         (1|word) + (1+context*vowel|subject), 
                         data= df_time, REML=FALSE)
summary(phon_acco.model)
summary(rePCA(phon_acco.model))

```

Power analysis:

```{r}

fixef(phon_acco.model)
powerSim(phon_acco.model)

trial <- lmer(f1 ~
                context*vowel + gender + task +
                (1+context*vowel|subject) + (1|word),
                data = df_time, REML = FALSE)

summary(trial)

fixef(trial)["contexte:tasks"] <- 0.1
powerSim(trial, test = fixed("tasks"), nsim = 20)
pc <- powerCurve (trial, along = "subject", test = fixed("task"), nsim = 20)
plot(pc)

null.model = lmer(f1 ~ 
                    gender + context*vowel + task + 
                    (1|word) + (1+context*vowel|subject), 
                    data=df_time, REML=FALSE)
summary(null.model)

anova(phon_acco.model,null.model)

coef(phon_acco.model)

```